---
title: "HW3"
author: "Yigao Li"
date: "September 23, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 17

```{r}
n = 100000
data17 = rbeta(n, 0.5, 0.8)
cat("E(X) =", mean(data17))
cat("\ns(X) =", sd(data17))
cat("\nE(X^(-1/3)) =", mean(data17^(-1/3)))
```

## Problem 18

```{r}
for (n in c(340000:350000)){
  data18 = rbeta(n, 0.5, 0.8)
  sample_mean = mean(sqrt(data18))
  sample_var = var(sqrt(data18))/n
  sample_sd = sqrt(sample_var)
  error = qnorm(0.975, mean = sample_mean, sd = sample_sd) - sample_mean
  if (error < 1e-3){
    sim = n
    break
  }
}
cat("Simulation of", sim, "times is necessary to approximate E(sqrt(x)) with an error of less than 10^-3.")
```

By Central Limit Theorem, when $n$ is as large as $340806$ times, we are 95% confident that this simulation to approximate $E[\sqrt{X}]$ with an error of less than $10^{-3}$.

## Problem 19

```{r}
while (TRUE){
  x = runif(1)
  y = runif(1)
  if (y <= (sin(2*pi*x))^2){
    Z = x
    break
  }
}

while (TRUE){
  x = runif(1)
  y = runif(1)
  if (y <= (sin(2*pi*x))^2){
    Z = append(Z, x)
    if (length(Z) == 10000){
      break
    }
  }
}

hist(Z)
```

Histogram shows reflective 2 peaks around 0.25 and 0.75. Function $y=sin^2(2\pi x)$ has a closely similar plot with this histogram.

## Problem 20

a)
Sensitivity is the probability that a drunk driver is correctly identified as drunk.
Specificity is the probability that a sober driver is correctly identified as being sober.

b) $Pr(Fail|Drunk)=0.99$, $Pr(Fail|Sober)=0.02$, $Pr(Drunk)=\frac{1}{500}$
$$Pr(Drunk|Fail)=\frac{Pr(Fail|Drunk)\times Pr(Drunk)}{Pr(Fail)}$$  
where
$$\begin{aligned}
Pr(Fail)&=Pr(Fail|Drunk)\times Pr(Drunk)+Pr(Fail|Sober)\times Pr(Sober)\\
&=Pr(Fail|Drunk)\times Pr(Drunk)+Pr(Fail|Sober)\times (1-Pr(Drunk))\\
&=0.99\times\frac{1}{500}+0.02\times\frac{499}{500}\\
&=0.02194\\
\end{aligned}$$
$$Pr(Drunk|Fail)=\frac{0.99\times\frac{1}{500}}{0.02194}\approx0.09024$$

## Problem 21

$X\sim Binom(n=50,p=0.2)$
(1)

```{r}
s20 = pbinom(19, 50, 0.2)
g10s20 = pbinom(19, 50, 0.2) - pbinom(10, 50, 0.2)
g10cs20 = g10s20/s20
cat("Pr(X<20) =", s20)
cat("\nPr(X>10|X<20) =", g10cs20)
```

(2) $X\sim Binom(n=500,p=0.1)$

(3)

```{r}
n = 1000
x = rbinom(n, 500, 0.1)
xs60 = sum(x<60)/n
xg30 = sum(x>30)/n
xg30s60 = sum((x>30)*(x<60))/n
xs60cg30 = xg30s60/xg30
xg30cs60 = xg30s60/xs60
cat("Pr(X<60) =", xs60)
cat("\nPr(X<60|X>30) =", xs60cg30)
cat("\nPr(X>30|X<60) =", xg30cs60)
```

## Problem 22

```{r}
size = 1000
n = 10
x = integer(size)
y = integer(size)
plot(x, y, xlim = c(-4,4), ylim = c(0,0.5))
for (i in c(1:n)){
  xsub = rcauchy(size)
  x = x + xsub
  xsub = sort(xsub)
  ysub = dcauchy(xsub)
  if (i == n){
    hist(x/n, freq = FALSE, breaks = 2000, add = TRUE)
  }
  lines(xsub, ysub, col = rainbow(x)[i])
}
```

## Problem 23

$$\begin{aligned}
Pr(Y\leq y)&=Pr[(X|X>A)-A\leq y]\\
&=Pr[(X|X>A)\leq A+y]\\
&=Pr(X\leq A+y|X>A)\\
&=1-Pr(X>A+y|X>A)\\
\end{aligned}$$
By Memoryless Property of Exponential Distribution,
$$\begin{aligned}
Pr(Y\leq y)&=1-Pr(X>y)\\
&=1-e^{-\lambda y}
\end{aligned}$$