---
title: "HW4"
author: "Yigao Li"
date: "October 1, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 25

```{r}
v <- c()
for (i in c(1:100)){
  x <- 0
  count <- 0
  while (TRUE){
    if (runif(1) > 0.5){
      x <- x + 1
    } else{
      x <- x - 1
    }
    count <- count + 1
    if (x == 15){
      break
    }
  }
  v = append(v, count)
}
boxplot(v)
summary(v)
```

## Problem 26

St. Petersburg strategy betting on a block of $6$ numbers.  
- Bet 1 dollar  
- After a loss, six times your last bet  
- After a win, bet 1 dollar again  
- Stop when you have lost all your money

## Problem 27

Let $X$ be the amount of dollars won or lost after a single game.  
$Pr(X=3)=\frac{10}{38}=\frac{5}{19}$  
$Pr(X=1)=\frac{8}{38}=\frac{4}{19}$  
$Pr(X=-1)=\frac{8}{38}=\frac{4}{19}$  
$Pr(X=-3)=\frac{12}{38}=\frac{6}{19}$  
$E[X]=3\times\frac{5}{19}+\frac{4}{19}-\frac{4}{19}-3\times\frac{6}{19}=-\frac{3}{19}$

## Problem 28

False. Assume $j$ location is one of the corner tiles and $k$ is the adjacent tile along the longer side. The probability that the frog will jump from $j$ to $k$ is $\frac{1}{2}$ but the probability from $k$ to $j$ is $\frac{1}{3}$. Therefore, this random walk is not symmetric.

## Problem 29

$X$ follows discrete uniform distribution $X\sim U(1,n)$ and $Y$ follows binomial distribution $Y\sim Binom(k,p)$.  
$$Pr(X=k)=\frac{1}{n}$$
$$Pr(Y=y|X=k)=\begin{pmatrix}k\\y\end{pmatrix}p^y(1-p)^{1-y}$$
$$\begin{aligned}
Pr(X=k,Y=y)&=Pr(Y=y|X=k)Pr(X=k)\\
&=\frac{1}{n}\begin{pmatrix}k\\y\end{pmatrix}p^y(1-p)^{1-y}\\
\end{aligned}$$

## Problem 30

a)
``` {r}
m <- c(1/2, 1/4, 0, 1/4, 0, 0, 0, 0, 0,
       1/6, 1/2, 1/6, 0, 1/6, 0, 0, 0, 0,
       0, 1/4, 1/2, 0, 0, 1/4, 0, 0, 0,
       1/6, 0, 0, 1/2, 1/6, 0, 1/6, 0, 0,
       0, 1/6, 0, 1/6, 1/2, 1/6, 0, 0, 0,
       0, 0, 1/4, 0, 1/4, 1/2, 0, 0, 0,
       0, 0, 0, 1/4, 0, 0, 1/2, 1/4, 0,
       0, 0, 0, 0, 0, 0, 1/4, 1/2, 1/4,
       0, 0, 0, 0, 0, 0, 0, 0, 1)
zog_transition_matrix <- matrix(m, ncol=9, nrow=9, byrow=T)
zog_matrix <- zog_transition_matrix %*% zog_transition_matrix %*% zog_transition_matrix
cat("P{ X3 = 3 | X0 = 1 } =", zog_matrix[1,3])
```

$$X_0=1
\begin{cases}
X_1=1(p=\frac{1}{2})
  \begin{cases}
  X_2=1(p=\frac{1}{2})\\
  X_2=2(p=\frac{1}{4})\implies X_3=3(p=\frac{1}{6})\\
  X_3=4(p=\frac{1}{4})\\
  \end{cases}
\\
X_1=2(p=\frac{1}{4})
  \begin{cases}
  X_2=1(p=\frac{1}{6})\\
  X_2=2(p=\frac{1}{2})\implies X_3=3(p=\frac{1}{6})\\
  X_2=3(p=\frac{1}{6})\implies X_3=3(p=\frac{1}{2})\\
  X_2=5(p=\frac{1}{6})\\
  \end{cases}
\\
X_1=4(p=\frac{1}{4})\\
\end{cases}$$
$$Pr(X_3=3|x_0=1)=\frac{1}{2}\times\frac{1}{4}\times\frac{1}{6}+\frac{1}{4}\times\frac{1}{2}\times\frac{1}{6}+\frac{1}{4}\times\frac{1}{6}\times\frac{1}{2}=\frac{1}{16}=0.0625$$

b)
``` {r}
m <- c(1/2, 1/4, 0, 1/4, 0, 0, 0, 0, 0,
       1/6, 1/2, 1/6, 0, 1/6, 0, 0, 0, 0,
       0, 1/4, 1/2, 0, 0, 1/4, 0, 0, 0,
       1/6, 0, 0, 1/2, 1/6, 0, 1/6, 0, 0,
       0, 1/6, 0, 1/6, 1/2, 1/6, 0, 0, 0,
       0, 0, 1/4, 0, 1/4, 1/2, 0, 0, 0,
       0, 0, 0, 1/4, 0, 0, 1/2, 1/4, 0,
       0, 0, 0, 0, 0, 0, 1/4, 1/2, 1/4,
       0, 0, 0, 0, 0, 0, 0, 0, 1)
zog_transition_matrix <- matrix(m, ncol=9, nrow=9, byrow=T)

a <- T
count <- 0
while (a){
  zog_last_matrix <- zog_transition_matrix
  zog_transition_matrix <- zog_transition_matrix %*% zog_transition_matrix
  count <- count + 1
  for (i in c(1:8)){
    a <- F
    if (zog_transition_matrix[i,9] < 0.5){
      a <- T
      break
    }
  }
}

b <- T
while (b){
  for (i in c(1:8)){
    if (zog_last_matrix[i,9] < 0.5){
      b <- F
    }
  }
  if (b == T){
    zog_last_matrix <- zog_transition_matrix
    zog_transition_matrix <- zog_transition_matrix %*% zog_transition_matrix
    count <- count + 1
  }
}

cat("The chance of the caveman's survival for more than", count,
    "steps is less 50% no matter where he starts.")
```

## Problem 31

a)
$N\sim Pois(\lambda)$, $Pr(N=n)=\frac{\lambda^ne^{-\lambda}}{n!}$  
$X\sim Binom(n,p)$, $Pr(X=k|N=n)=\begin{pmatrix}n\\k\end{pmatrix}p^k(1-p)^{n-k}$  
$Pr(X=k,N=n)=\frac{\lambda^ne^{-\lambda}}{n!}\begin{pmatrix}n\\k\end{pmatrix}p^k(1-p)^{n-k}$

b)
```{r}
simX <- function(lambda, p, k){
  n <- rpois(k, lambda = lambda)
  b <- rbinom(k, size = n, prob = p)
}
```

c)
$\lambda=44$, $p=0.5$, $E[X]=\lambda p$
```{r}
mean(replicate(100, simX(44, 0.5, 100)))
```

## Problem 32

a)
$Pr(X<12|X<18)=\frac{Pr(X<12,X<18)}{Pr(X<18)}=\frac{Pr(X<12)}{Pr(X<18)}$
```{r}
pbinom(11, 80, 0.2)
pbinom(17, 80, 0.2)
pbinom(11, 80, 0.2)/pbinom(17, 80, 0.2)
```
$Pr(X<12|X<18)\approx0.1499783$

$E[X|X<18]$
```{r}
E <- 0
for (i in c(0:17)){
  E <- E + i*dbinom(i, 80, 0.2)/pbinom(17, 80, 0.2)
}
print(E)
```
$E[X|X<18]\approx14.0393$

b)
```{r}
x = 12:20
denominator = pbinom(20, 80, 0.2) - pbinom(11, 80, 0.2)
y = sapply(x, function (x) (pbinom(x, 80, 0.2) - pbinom(11, 80, 0.2))/denominator)
plot(x, y, type = 'b', main = 'CDF of X|(12<=X<=20)', ylab = 'f(x|(12<=X<=20))')
```

c)
```{r}
n = 1e4
xcz <- c()
x <- rbinom(n, 80, 0.2)
y <- rbinom(n, 100, 0.7)
for (i in c(1:n)){
  if (x[i] + y[i] == 90){
    xcz = c(xcz, x[i])
  }
}
plot(ecdf(xcz), main = 'CDF of X|Z=90', ylab = 'f(x|z=90)')
```

d)
```{r}
n = 1e4
x <- rbinom(n, 80, 0.2)
y <- rbinom(n, 100, 0.7)
z <- c()
k <- c(10, 15, 20)
for (kvalue in k){
  for (i in c(1:n)){
    if (x[i] == kvalue){
      z <- c(z, x[i]+y[i])
    }
  }
  cat("E[Z|X=", kvalue, "]=", mean(z), "\n")
}
```

e)
```{r}
n = 1e4
x <- rbinom(n, 80, 0.2)
y <- rbinom(n, 100, 0.7)
xczk <- c()
k <- c(80, 90, 100)
for (kvalue in k){
  for (i in c(1:n)){
    if (x[i]+y[i] == kvalue){
      xczk <- c(xczk, x[i])
    }
  }
  cat("E[X|Z=", kvalue, "]=", mean(xczk), "\n")
}
```