---
title: "HW6"
author: "Yigao Li"
date: "October 29, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}
n <- 1000
z <- replicate(n, sum(runif(12))-6)
qqnorm(z)
```

$X_i\sim Uniform(0,1)$, $E[Z]=E[X_1+X_2+\dots+X_{12}-6]=E[X_1]+E[X_2]+\dots+E[X_{12}]-6=\frac12\times12-6=0$, $Var(Z)=Var(X_1+X_2+\dots+X_{12}-6)=Var(X_1)+Var(X_2)+\dots+Var(X_{12})=\frac1{12}\times12=1$

## Problem 2

(a) Expected value of a sample mean is $\frac1{\lambda}=10$

(b)

```{r}
n <- 1000
samples <- replicate(n, mean(rexp(30, 1/10)))
print(sum(samples >= 12)/n)
```

(c) The probability of the sample mean no less than 12 is greater than 0.05. There is not enough evidence that a mean of 12 is unusual.

## Problem 3

(a)

```{r}
phillies <- read.csv("D:/Courses/ANLY 511/Phillies2009.csv")
strikeoutsHome <- subset(phillies$StrikeOuts, phillies$Location == "Home")
strikeoutsAway <- subset(phillies$StrikeOuts, phillies$Location == "Away")
plot(ecdf(strikeoutsHome), main = "Empirical Distribution Function of StrikeOuts at Home",
     xlab = "StrikeOuts", xlim = c(0,20))
plot(ecdf(strikeoutsAway), main = "Empirical Distribution Function of StrikeOuts at Away",
     xlab = "StrikeOuts", xlim = c(0,20))
```

(b)

```{r}
aggregate(StrikeOuts ~ Location, data = phillies, FUN = mean)
```

(c)

$H_0:\mu_{StrikeOutsHome}=\mu_{StrikeOutsAway}$  
$H_1:\mu_{StrikeOutsHome}<\mu_{StrikeOutsAway}$

```{r}
mytest.1 <- function(mydf){
  agg <- aggregate(StrikeOuts ~ Location, data = mydf, FUN = mean)
  return(agg$StrikeOuts[1] - agg$StrikeOuts[2])
}

permute.sample.1 <- function(mydf){
  n <- dim(mydf)[1]
  mydf$Location <- mydf$Location[sample(n, n, replace = F)]
  return(mytest.1(mydf))
}

phillies.permute <- phillies
N <- 1000
test.1 <- replicate(N, permute.sample.1(phillies.permute))
hist(test.1, main = "Null Distribution", prob = T,
     sub = paste("p-value =", mean(test.1 > mytest.1(phillies))))
abline(v = mytest.1(phillies), col = 2, lwd = 2)
```

Because p-value is greater than $0.05$, we fail to reject $H_0$ at the significance level $0.05$.

## Problem 4

$H_0$: winning numbers are randomly drawn.  
$H_1$: winning numbers are not randomly drawn.
```{r}
lottery <- read.csv("D:/Courses/ANLY 511/Lottery.csv")
myX2 <- function(x, n, pvec){
  k <- length(pvec)
  pvec <- pvec/sum(pvec)
  expected <- n*pvec
  return(sum((x - expected)^2/expected))
}
lotterytable <- table(lottery)
freq <- as.vector(lotterytable)
mySim.1 <- function(n, pvec){
  x <- rmultinom(1, n, pvec)
  x <- as.vector(x)
  return(myX2(x, n, pvec))
}
myX2test <- function(x.obs, n, pvec, N){
  myX2dist <- replicate(N, mySim.1(n, pvec))
  X2.obs <- myX2(x.obs, n, pvec)
  p.value <- mean(myX2dist > X2.obs)
  hist(myX2dist, main = "Null Distribution", sub = paste("p-value =", p.value), xlab = "X2", prob = T)
  abline(v = X2.obs, col = 2, lwd = 2)
  return(p.value = p.value)
}
myX2test(freq, 500, rep(1,39), 10000)
```

Since p-value is greater than 0.05. We fail to reject $H_0$ at significance level $0.05$.

## Problem 5

```{r}
samplef <- function(x){var(rnorm(20, 25, 7))}
my.vars <- numeric(1000)
my.vars <- sapply(my.vars, samplef)
mean(my.vars)
var(my.vars)
hist(my.vars)
qqnorm(my.vars)
qqline(my.vars)

samplef <- function(x){var(rnorm(50, 25, 7))}
my.vars <- numeric(1000)
my.vars <- sapply(my.vars, samplef)
mean(my.vars)
var(my.vars)
hist(my.vars)
qqnorm(my.vars)
qqline(my.vars)

samplef <- function(x){var(rnorm(200, 25, 7))}
my.vars <- numeric(1000)
my.vars <- sapply(my.vars, samplef)
mean(my.vars)
var(my.vars)
hist(my.vars)
qqnorm(my.vars)
qqline(my.vars)
```

Sampling distribution appears to be normally distributed when $n=200$.

## Problem 6

(a) $W=\bar X-\bar Y=\frac19X_1+\frac19X_2+\dots+\frac19X_9-\frac1{12}Y_1-\frac1{12}Y_2-\dots-\frac1{12}Y_{12}$.  
Because $W$ is a linear combination of normal distributions $X_i, Y_i$, $W$ is also a normal distribution with mean $E[W]=\frac19E[X_1]+\frac19E[X_2]+\dots+\frac19E[X_9]-\frac1{12}E[Y_1]-\frac1{12}E[Y_2]-\dots-\frac1{12}E[Y_{12}]=7-10=-3$, and variance $Var(W)=\frac1{81}Var(X_1)+\frac1{81}Var(X_2)+\dots+\frac1{81}Var(X_9)+\frac1{144}Var(Y_1)+\frac1{144}Var(Y_2)+\dots+\frac1{144}Var(Y_{12})=\frac19\times3^2+\frac1{12}\times5^2=\frac{37}{12}$

(b)

```{r}
W <- numeric(1000)
for (i in 1:1000){
  x <- rnorm(9, 7, 3)
  y <- rnorm(12, 10, 5)
  W[i] <- mean(x) - mean(y)
}
hist(W, prob = T)
```

(c)

```{r}
mean(W< (-1.5))
pnorm(-1.5, -3, sqrt(37/12))
```

## Problem 7

(a) W is approximately normally distributed with mean $E[W]=\frac19E[X_1]+\frac19E[X_2]+\dots+\frac19E[X_9]-\frac1{12}E[Y_1]-\frac1{12}E[Y_2]-\dots-\frac1{12}E[Y_{12}]=0.5-1=-0.5$, and variance $Var(W)=\frac1{81}Var(X_1)+\frac1{81}Var(X_2)+\dots+\frac1{81}Var(X_9)+\frac1{144}Var(Y_1)+\frac1{144}Var(Y_2)+\dots+\frac1{144}Var(Y_{12})=\frac19\times\frac1{12}+\frac1{12}\times\frac1{12}=\frac7{432}$

(b)

```{r}
W <- numeric(1000)
for (i in 1:1000){
  x <- runif(9)
  y <- runif(12, 0.5, 1.5)
  W[i] <- mean(x) - mean(y)
}
hist(W, prob = T)
```

(c)
(Note: this part maybe is wrong because the maximum possible value for W is 0.5, which is always less than 0.6)
```{r}
mean(W< 0.6)
pnorm(0.6, -0.5, sqrt(7/432))
```

## Problem 8

```{r}
titani <- read.csv("D:/Courses/ANLY 511/Titanic.csv")
mytest.2 <- function(mydf){
  agg <- aggregate(Age ~ Survived, data = mydf, FUN = mean)
  return(agg$Age[1] - agg$Age[2])
}

permute.sample.2 <- function(mydf){
  n <- dim(mydf)[1]
  mydf$Survived <- mydf$Survived[sample(n, n, replace = F)]
  return(mytest.2(mydf))
}

titani.permute <- titani
N <- 1000
test.2 <- replicate(N, permute.sample.2(titani.permute))
hist(test.2, main = "Null Distribution", prob = T,
     sub = paste("p-value =", mean(abs(test.2) > mytest.2(titani))), xlim = c(-5,5))
abline(v = mytest.2(titani), col = 2, lwd = 2)
abline(v = -mytest.2(titani), col = 2, lwd = 2)
```

Because p-value is smaller than $0.05$, we reject $H_0$ at the significance level $0.05$.