---
title: "HW9"
author: "Yigao Li"
date: "November 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 (7-9)
(a)
```{r}
spruce <- read.csv("D:/Courses/ANLY 511/Spruce.csv")
hist(spruce$Ht.change, prob = TRUE, main = "Distribution of Ht.change")
```
(b)
```{r}
dHeight <- spruce$Height5 - spruce$Height0
n <- length(dHeight)
dHeight.s <- sd(dHeight)
dHeight.mu <- mean(dHeight)
error <- dHeight.s/sqrt(n)
talpha <- qt(0.975, n-1)
cat("95% t confidence interval for the mean height change over the 5-year period is [", dHeight.mu-talpha*error, ",", dHeight.mu+talpha*error, "].")
```
We are 95% confident that the true height change is between 28.33685 and 33.52982.

## Problem 2 (7-10)

```{r}
n1 <- 43
x1 <- 5.8
s1 <- 8.6
n2 <- 36
x2 <- 1.9
s2 <- 4.2
talpha <- qt(0.975, n1+n2-2)
me <- talpha * sqrt(s1^2/n1 + s2^2/n2)
cat("95% t confidence interval for the mean difference in weight loss between 2 groups is [", x1-x2-me, ",", x1-x2+me, "].")
```
We are 95% confident that the true difference in weight loss between 2 groups is between 0.9397875 and 6.860212.

## Problem 3 (7-20)
(a)
```{r}
nf <- 980
mf <- 459
nm <- 759
mm <- 426
pf_hat <- mf/nf
zalpha <- qnorm(0.975)
lowerbound <- pf_hat - zalpha * sqrt(pf_hat*(1-pf_hat)/nf)
upperbound <- pf_hat + zalpha * sqrt(pf_hat*(1-pf_hat)/nf)
cat("95% confidence interval for the proportion of women who voted for Bush is [", lowerbound, ",", upperbound, "].")
```
(b)
```{r}
pm_hat <- mm/nm
lowerbound <- pm_hat - zalpha * sqrt(pm_hat*(1-pm_hat)/nm)
upperbound <- pm_hat + zalpha * sqrt(pm_hat*(1-pm_hat)/nm)
cat("95% confidence interval for the proportion of men who voted for Bush is [", lowerbound, ",", upperbound, "].")
```
No, confidence intervals for men and women do not overlap. The proportion of men voting for Bush is higher than that of women.  
(c)
```{r}
p_hat <- (mf+mm)/(nf+nm)
se <- sqrt(p_hat*(1-p_hat)*(1/nf+1/nm))
lowerbound <- pm_hat - pf_hat - zalpha * se
upperbound <- pm_hat - pf_hat + zalpha * se
cat("95% confidence interval for the difference in proportions (men - women) is [", lowerbound, ",", upperbound, "].")
```
We are 95% confident that proportion of men voting for Bush is 4.552076% to 14.02742% higher than proportion of women voting for Bush.

## Problem 4 (8-3)
(a)  
$H_0$: $\mu=98.6$  
$H_a$: $\mu>98.6$  

(b)  
```{r}
temperature <- c(98, 98.9, 99, 98.9, 98.8, 98.6, 99.1, 98.9, 98.5, 98.9, 98.9, 98.4, 99, 99.2, 98.6, 98.8, 98.9, 98.7)
t.test(temperature, alternative = "greater", mu = 98.6)
```
P-value is $0.006907<\alpha=0.05$. We reject $H_0$ and conclude that body temperatures in children in Sober are higher than the standard one.

## Problem 5 (8-10)
$H_0$: $\hat p=29.1\%$  
$H_a$: $\hat p<29.1\%$
```{r}
prop.test(87, 350, p = 0.291, alternative = "less")
```
P-value is $0.04564<\alpha=0.05$. We reject $H_0$ and sample data supports professor's hypothesis.

## Problem 6 (7-30)
Without loss of generality, assume that $\hat\theta_1\geq\hat\theta_2$.  
$(\implies)\hat\theta_1\pm1.96\hat{SE}_1$ and $\hat\theta_2\pm1.96\hat{SE}_2$ overlap. Therefore,  
$$\hat\theta_1-1.96\hat{SE}_1\leq\hat\theta_2+1.96\hat{SE}_2$$
$$\hat\theta_1-\hat\theta_2-1.96(\hat{SE}_1+\hat{SE}_2)\leq0$$
Since $\hat\theta_1-\hat\theta_2\geq0$, $\hat{SE}_1,\hat{SE}_2\geq0$,  
$\hat\theta_1-\hat\theta_2+1.96(\hat{SE}_1+\hat{SE}_2)\geq0$$
Therefore, $\hat\theta_1-\hat\theta_2\pm1.96(\hat{SE}_1+\hat{SE}_2)$ must contain $0$.  
$(\impliedby)\hat\theta_1-\hat\theta_2\pm1.96(\hat{SE}_1+\hat{SE}_2)$ contains $0$  
$$\hat\theta_1-\hat\theta_2+1.96(\hat{SE}_1+\hat{SE}_2)\geq0$$
$$\hat\theta_1-\hat\theta_2-1.96(\hat{SE}_1+\hat{SE}_2)\leq0$$
So,  
$$\hat\theta_1-1.96\hat{SE}_1\leq\hat\theta_2+1.96\hat{SE}_2$$
Because $\hat\theta_1\geq\hat\theta_2$, $\hat\theta_1\pm1.96\hat{SE}_1$ and $\hat\theta_2\pm1.96\hat{SE}_2$ must overlap.

## Problem 7 (7-33)
(a)  
Let $s^2$ be sample variance. By Central Limit Theorem, $\lambda\sim N(\bar X, \frac{s^2}n)$. Therefore, 95% confidence interval is [$\bar X-1.96\frac s{\sqrt{n}},\bar X+1.96\frac s{\sqrt{n}}$]  
(b)  
```{r}
x <- c(4,6,7,9,10,13)
mean(x)
var(x)
```
$\bar X=\frac{49}6$, $s^2=\frac{61}6$, $n=6$, 95% confidence interval [$5.6,10.7$]  

## Problem 8 (8-22)
(a)  
```{r}
ppois(16/5, 2, lower.tail = FALSE)
```
The probability of a Type I error is 14.28765%.  
(b)  
```{r}
ppois(16/5, 4, lower.tail = TRUE)
```
The probability of a Type II error is 43.34701%.